---
title: "Common barriers to Covid-19 Vaccination"
author: Naijia Liu
date: 2021-04-23
tags:
  - COVIDcast
  - symptom surveys
  - vaccine
authors:
  - alex
  - robin
  - naijia
heroImage: 
heroImageThumb: 
related:
  - 
summary: |
  Common barriers to vaccination by vaccination status and 
  different age groups.
acknowledgements: |

output:
  html_document:
    code_folding: hide
  blogdown::html_page:
    toc: true
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## This post's filename starts with _ so it is not run by blogdown during the
## post build process, since it depends on private microdata. Instead, we commit
## the HTML.
library(knitr)
library(ComplexUpset)
library(qdapTools)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(gt)
library(xts)
library(directlabels)
library(ggrepel)
library(data.table)
library(readr)
library(scales)

knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## data we will be using for the post

# load the data we need for Wave 11 in May
survey5_full <- fread("/Users/bb/Box/CTIS\ data/2021-05-race-ethnicity.csv.gz")


# remove barrier 11 "None of the above"
survey5 <- fread("/Users/bb/Box/CTIS\ data/2021-05-race-ethnicity.csv.gz") %>% 
  filter(is.na(V15a) | V15a != "11",
         is.na(V15b) | V15b != "11")

survey5$barriers_logic <- split_options(coalesce(survey5$V15a, survey5$V15b))


# load the data we need for trend plot
# read in the data
survey_month <- fread("/Users/bb/Box/CTIS\ data/monthly_nation_age.csv.gz")


# load the data we need for Wave 12 in Jan,2022
survey1_full <- fread("/Users/bb/Box/CTIS\ data/2022-01-race-ethnicity.csv.gz")

# new barriers were introduced in Wave 12 in Jan,2022
survey1 <- fread("/Users/bb/Box/CTIS\ data/2022-01-race-ethnicity.csv.gz") %>% 
  filter(is.na(V15c) | V15c != "11",
         is.na(V15b) | V15b != "11")

survey1$barriers_logic <- split_options(coalesce(as.character(survey1$V15c), survey1$V15b))

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load the fuctions we need
split_options <- function(column) {
  return(strsplit(column, ",", fixed = TRUE))
}

is_selected <- function(vec, selection) {
  selections <- unlist(lapply(
    vec,
    function(resp) {
      if (length(resp) == 0 || all(is.na(resp))) {
        # Qualtrics files code no selection as "" (empty string), which is
        # parsed by `read_csv` as `NA` (missing) by default. Since all our
        # selection items include "None of the above" or similar, treat both no
        # selection ("") or missing (NA) as missing, for generality.
        NA
      } else {
        selection %in% resp
      }
    }))
  return(selections)
}

code_demographics <- function(d) {
  d$race_ethnicity <- case_when(
    d$D6 == 1 ~ "Hispanic",
    d$D7 == 3 ~ "Black",
    d$D7 == 5 ~ "White",
    is.na(d$D7) | is.na(d$D7) ~ NA_character_,
    TRUE ~ "Other"
  )}
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Now let's code barriers for May,2021
barriers <- list(
  "1" = "Eligibility requirements",
  "2" = "No vaccines or appointments",
  "3" = "Times didn't work",
  "4" = "Technical difficulties",
  "5" = "Document requirements",
  "6" = "Limited Internet/phone access",
  "7" = "Can't travel to vaccine site",
  "8" = "Language barrier",
  "9" = "Childcare",
  "10" = "Can't get time off",
  "12" = "Desired brand not available",
  "13" = "Have not tried yet"
)

survey5$barriers <- lapply(
  survey5$barriers_logic,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers[s])
    }
  }
)

# extract barrier string from barriers column
survey5$barriers_str <- sapply(
  survey5$barriers,
  function(ss) {
    paste0(ss, collapse = "&")
  }
)

# add column for vaccination status
survey5$vaccinated <- case_when(
  survey5$V1 == 1 ~ "Vaccinated",
  survey5$V1 == 2 ~ "Not vaccinated"
)


# regroup the age into 3 main groups
# "18-34 years", "35-64 years", "65 years or older"
age_logic <- list(
  "1" = "18-24 years",
  "2" = "25-44 years",
  "3" = "25-44 years",
  "4" = "45-64 years",
  "5" = "45-64 years",
  "6" = "65 years or older",
  "7" = "65 years or older")


## extract the subset for May v15a values
# one hot coding for the May v15a values
v15a_input <- mtabulate(survey5$barriers[which(survey5$V1 == 1)])

# add "age" column to the v15a data frame 
v15a_input$age <- lapply(
  as.character(survey5$D2[which(survey5$V1 == 1)]), # for those who are vaccinated
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(age_logic[s])
    }
  }
)

v15a_input$age <- as.character(v15a_input$age)

#remove the "NA" rows for vaccinated
v15a_input <- v15a_input[-which(is.na((survey5$barriers[which(survey5$V1 == 1)]))),]



## extract the subset for May v15b values
# one hot coding for the v15b values
v15b_input <- mtabulate(survey5$barriers[which(survey5$V1 == 2)])

# add "age" column to the v15b data frame 
v15b_input$age <- lapply(
  as.character(survey5$D2[which(survey5$V1 == 2)]),
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(age_logic[s])
    }
  }
)

v15b_input$age <- as.character(v15b_input$age)

#remove the "NA" rows (the highest pike in the plot) for not vaccinated
v15b_input <- v15b_input[-which(is.na((survey5$barriers[which(survey5$V1 == 2)]))),]
```

```{r}
## trend plot for all the vaccinated
# exclude "none" barrier
month_barrier_vaccinated <- data.frame(as.numeric(survey_month$period_val),
                      survey_month$age,
                      survey_month$val_pct_vaccine_barrier_eligible_has,
                      survey_month$val_pct_vaccine_barrier_no_appointments_has,
                      survey_month$val_pct_vaccine_barrier_appointment_time_has,
                      survey_month$val_pct_vaccine_barrier_technical_difficulties_has,
                      survey_month$val_pct_vaccine_barrier_document_has,
                      survey_month$val_pct_vaccine_barrier_technology_access_has,
                      survey_month$val_pct_vaccine_barrier_travel_has,
                      survey_month$val_pct_vaccine_barrier_language_has,
                      survey_month$val_pct_vaccine_barrier_childcare_has,
                      survey_month$val_pct_vaccine_barrier_time_has,
                      survey_month$val_pct_vaccine_barrier_type_has)

# only take the data from May,2021 to Jan,2021
month_barrier_vaccinated <- month_barrier_vaccinated[21:70,]

colnames(month_barrier_vaccinated) <- c("period_val","age","eligible","no_appointment",
                             "appointment_time","technical_difficultie",
                             "document","technology_access","travel","language",
                             "childcare","time","type")

# convert period into "year/month" format
date <- paste("2021", month_barrier_vaccinated$period_val[1:40], sep = "")
date <- append(date,paste("2022", month_barrier_vaccinated$period_val[41:50], sep = ""))

month_barrier_vaccinated$period_val <- as.yearmon(date,"%Y%m")

# aggregate the age groups for all vaccinated
month_barrier_vaccinated_all <- month_barrier_vaccinated %>% group_by(period_val) %>%
  summarise(eligible_all = sum(eligible),
            no_appointment_all = sum(no_appointment),
            appointment_time_all = sum(appointment_time),
            technical_difficultie_all = sum(technical_difficultie),
            document_all = sum(document),
            technology_access_all = sum(technology_access),
            travel_all = sum(travel),
            language_all = sum(language),
            childcare_all = sum(childcare),
            time_all = sum(time),
            type_all = sum(type))

colnames(month_barrier_vaccinated_all) <- colnames(month_barrier_vaccinated)[c(1,3:15)]

# convert the data frame into pivot long table
month_barrier_vaccinated_all_long <- month_barrier_vaccinated_all %>%
  pivot_longer(cols = eligible:type,
               names_to = "barriers",
               values_to = "percentage")

# Now let's code barriers
barriers_name <- list(
  "eligible" = "Eligibility requirements",
  "no_appointment" = "No vaccines or appointments",
  "appointment_time" = "Times didn't work",
  "technical_difficultie" = "Technical difficulties",
  "document" = "Document requirements",
  "technology_access" = "Limited Internet/phone access",
  "travel" = "Can't travel to vaccine site",
  "language" = "Language barrier",
  "childcare" = "Childcare",
  "time" = "Can't get time off",
  "type" = "Desired brand not available"
)

month_barrier_vaccinated_all_long$barriers <- lapply(
  month_barrier_vaccinated_all_long$barriers,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers_name[s])
    }
  }
)

# barriers column is a named list
# convert it into a list
barriers_new <- c()
for (i in 1:length(month_barrier_vaccinated_all_long$barriers)){
  barriers_new <- c(barriers_new,
                    (toString(unname(month_barrier_vaccinated_all_long$barriers)[i])))
}

month_barrier_vaccinated_all_long$barriers <- barriers_new
```


```{r}
## trend plot for the vaccinated by different age groups
# convert the data frame into pivot long table
month_barrier_vaccinated_long <- month_barrier_vaccinated %>%
  pivot_longer(cols = eligible:type,
               names_to = "barriers",
               values_to = "percentage")

month_barrier_vaccinated_long$barriers <- lapply(
  month_barrier_vaccinated_long$barriers,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers_name[s])
    }
  }
)

# barriers column is a named list
# convert it into a list
barriers_new <- c()
for (i in 1:length(month_barrier_vaccinated_long$barriers)){
  barriers_new <- c(barriers_new,
                    (toString(unname(month_barrier_vaccinated_long$barriers)[i])))
}

month_barrier_vaccinated_long$barriers <- barriers_new
```


```{r}
## prepare data for the unvaccinated table for Jan,2022
barriers_jan <- list(
  "1" = "Eligibility requirements",
  "2" = "No vaccines or appointments",
  "3" = "Times didn't work",
  "4" = "Technical difficulties",
  "5" = "Document requirements",
  "6" = "Limited Internet/phone access",
  "7" = "Can't travel to vaccine site",
  "8" = "Language barrier",
  "9" = "Childcare",
  "10" = "Can't get time off",
  "12" = "Desired brand not available",
  "13" = "Have not tried yet",
  "14" = "Available locations did't work",
  "15" = "Other"
)

# data wangling for barriers
survey1$barriers <- lapply(
  survey1$barriers_logic,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers_jan[s])
    }
  }
)

# extract barrier string from barriers column
survey1$barriers_str <- sapply(
  survey1$barriers,
  function(ss) {
    paste0(ss, collapse = "&")
  }
)

# add column for vaccination status
survey1$vaccinated <- case_when(
  survey1$V1 == 1 ~ "Vaccinated",
  survey1$V1 == 2 ~ "Not vaccinated"
)

# add "age" column to the data frame 
survey1$age <- lapply(
  survey1$D2,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(age_logic[s])
    }
  }
)

# extract the subset for Jan 2022 v15b values
# one hot coding for the Jan 2022 v15b values
jan_v15b_input <- survey1 %>% filter(V1==2 & !is.na(V15b) & V15b!=11) %>%
  select(barriers,age)

```


On May 10th, 2021,the Centers for Disease Control and Prevention (CDC) and 
the Food and Drug Administration
[expanded the emergency use authorization (EUA) for the Pfizer-BioNTech COVID-19 Vaccine](https://www.fda.gov/news-events/press-announcements/coronavirus-covid-19-update-fda-authorizes-pfizer-biontech-covid-19-vaccine-emergency-use) to include adolescents 12 through 15 
years of age, amending the EUA originally issued on Dec. 11, 2020 for 
administration in individuals 16 years of age and older. This declared the 
universal acceptance for vaccination, and another significant step to return to 
normality.

By understanding and assisting overcome the barriers to vaccination, one of the 
major goals for Delphi group has always been helping raise the vaccination rate 
and resume the normal life. And that’s why, through its COVID-19 Trends and 
Impact Survey, the Delphi Group started to track COVID vaccination barriers 
right after FDA’s announcement, to help understand the vaccination barriers by 
vaccination status and age groups. Delphi’s COVID-19 Trends and Impact Survey
is a massive survey distributed daily across the United States through our 
partnership with Facebook, and with the help of it, we’ve successfully tracked `r blogdown::shortcode_html("reflink", "2021-01-22-holiday-surveys", "social distancing")` 
and `r blogdown::shortcode_html("reflink", "2020-12-10-masks-public", "mask use")`,
and gave some thoroughly insights of `r blogdown::shortcode_html("reflink", "2021-04-20-jj-vaccine", "vaccine hesitancy")`.

Since May 20th, 2021, we have asked all respondents about whether they have 
experienced specific barriers to getting the COVID-19 vaccine based on
their vaccination status. Combined with their answers on age groups, we can have
a general idea of the common barriers to vaccination by vaccination status and 
age groups, and make monthly comparison from May 2022 to now.

From our finding, based on the different vaccination status and age groups, our 
survey respondents have encountered different barriers to vaccination. Though, 
our data does suggest that extra efforts should be taken to ensure enough 
availability to vaccination for potential vaccine recipients.

Not surprisingly, CDC has already suggested multiple approaches that can 
[help increase COVID-19 Vaccination](https://www.cdc.gov/vaccines/covid-19/health-departments/generate-vaccinations.html#:~:text=Partner%20with%20local%20community%20groups,or%20request%20at%2Dhome%20vaccinations),
including reaching people where they live, work, learn, pray, play, gather or 
through existing programs and the internet. For example, CDC mentions providing 
at-home vaccination for those have difficulty accessing a vaccination, like older
adults, is a great way to increase vaccine rate for priority populations. 
Besides, by offering options including allowing employees to be away during work
hours or to taking paid leave to get vaccinated, reimbursing employees for 
transportation costs, and offering on-site vaccination, employers can help 
employees get vaccinated. We hope, through this article, we can give more 
insightful suggestions on how we could make vaccines more available to more 
potential recipient by different age groups.


Let’s start with the common barriers to vaccination on May, 2021 by vaccination 
status. First, we ask respondents whether they have been vaccinated or not, and
according to their vaccination status, they will be given a list of 12 or 13 
specific barriers. Respondents can select multiple reasons, so the numbers do 
not add to 100%. One thing we also need to notice is the huge sample size 
disparity between the vaccinated and the unvaccinated, which could possibly 
affect the accuracy of the measurement for all respondents.


```{r}
# percentage table for different barriers by vaccination status on May,2021

## sample size without removing "none of above" (barrier 11)
# sample size for the vaccinated May
samplesize_v5 <- survey5_full %>% filter(V1==1 & !is.na(V15a)) %>% nrow
# sample size for the unvaccinated May
samplesize_unv5 <- survey5_full %>% filter(V1==2 & !is.na(V15b)) %>% nrow
# sample size for all May
samplesize_all5 <- samplesize_v5+samplesize_unv5

## sample size removing "none of above" (barrier 11)
## to match with the trend plot
# sample size for the vaccinated May
samplesize_v5_use <- survey5_full %>% filter(V1==1 & !is.na(V15a) & V15a!=11) %>% nrow
# sample size for the unvaccinated May
samplesize_unv5_use <- survey5_full %>% filter(V1==2 & !is.na(V15b) & V15b!=11) %>% nrow
# sample size for all May
samplesize_all5_use <- samplesize_v5_use+samplesize_unv5_use

# percentage for barriers who have vaccinated
freq_vaccinated <- table(unlist(survey5$barriers[which(survey5$V1 == 1)]))
prop_vaccinated <- as.data.frame(freq_vaccinated/samplesize_v5_use)

vac <- as.double(pull(prop_vaccinated, Freq))
vac <- c(vac[1:6],0,vac[7:11]) # no barrier "have not tried yet" for the vaccinated

# percentage for barriers who have not vaccinated
freq_not_vaccinated <- table(unlist(survey5$barriers[which(survey5$V1 == 2)]))
prop_not_vaccinated <- as.data.frame(freq_not_vaccinated/samplesize_unv5_use)

nvac <- pull(prop_not_vaccinated, Freq)

# percentage for barriers for all
freq <- table(unlist(survey5$barriers))
prop <- as.data.frame(freq/(samplesize_v5_use+samplesize_unv5_use))

table <- mutate(prop, vac, nvac)
table <- table[with(table,order(Freq,decreasing = TRUE)),]

# order the table by frequency of barriers of all vaccinated in descending order
table <- table[with(table,order(Freq,decreasing = TRUE)),]
table[,2:4] <- apply(table[,2:4], 2, function(x) percent(x, accuracy=0.01))

colnames(table) <- c("Barriers","All", "Vaccinated","Unvaccinated")

# transform the barriers abbreviation into long ones
barrier_explain <- list(
  "Eligibility requirements"="I did not meet the eligibility requirements",
  "No vaccines or appointments"="There were no vaccines or vaccine appointments available",
  "Times didn't work"="The available appointment times did not work for me",
  "Technical difficulties"="There were technical difficulties with the website or phone line",
  "Document requirements"="I was unable to provide a required document",
  "Limited Internet/phone access"="Limited access to internet or phone to schedule an appointment",
  "Can't travel to vaccine site"="Difficulty traveling to a vaccination site",
  "Language barrier"="Information not available in my native language",
  "Childcare"="There is no one to provide childcare while getting the vaccine",
  "Can't get time off"="It was difficult to get time away from work or school",
  "Desired brand not available"="I could not get the type of vaccine I wanted",
  "Have not tried yet"="I have not tried to get the vaccine",
  "Available locations did't work"="The available appointment locations did not work for me",
  "Other"="Other")

table$Barriers <- lapply(table$Barriers, function(s) unlist(barrier_explain[s]))

```


```{r}
# comparison table on May, 2021 by vaccination status
gt(table) %>%
  cols_align(
    align = "left",
    columns = Barriers
  ) %>%
    cols_align(
    align = "right",
    columns = c(All, Vaccinated, Unvaccinated)
  ) %>%
  tab_header(
    title = md("**Common barriers to vaccination**"),
    subtitle = md("*May, 2021*")
  ) %>%
    tab_footnote(paste0("Vaccinated sample size, n = ", 
                        prettyNum(samplesize_v5, big.mark = ",")),
                 cells_column_labels("Vaccinated")) %>%
    tab_footnote(paste0("Unvaccinated sample size, n = ", 
                        prettyNum(samplesize_unv5, big.mark = ",")),
                 cells_column_labels("Unvaccinated")) %>%
    tab_footnote(paste0("All sample size, n = ", 
                        prettyNum(samplesize_all5, big.mark = ",")),
                 cells_column_labels("All"))
```

As we can see, there are similarities and differences between the common barriers 
to vaccination for the vaccinated and the unvaccinated. "No vaccines or vaccine 
appointments" is the top selected barrier for both. Nevertheless, the other two 
top reasons commonly selected by the vaccinated are eligibility requirements 
and technical difficulties, while the unvaccinated select time didn’t work 
and desired brand not available most, which is unexpected for us. What is 
noticeable is that  barrier "nave not tried yet" also stands high among the 
unvaccinated. This is what we would expect to happen in normal circumstances: 
As more people get vaccinated, the people who remain unvaccinated are those who 
either don’t want to be vaccinated or haven’t yet been able to.

But these numbers don’t tell the whole story. It’s also very insightful to look 
at the time trend plot for barriers. The common barriers to vaccination varied 
throughout the past half year with the announcement for boosters, new policies, 
the larger population of the vaccinated and so on. In order to have clearer 
overview on the changes, we plot the trend for the barriers to vaccination for 
the vaccinated from May, 2021 to February, 2022, and two major discoveries are s
potted:


```{r}
# specify the factor levels for barriers in the descending order
month_barrier_vaccinated_all_long$barriers <- factor(
  month_barrier_vaccinated_all_long$barriers, 
  levels = c("No vaccines or appointments", "Times didn't work",
             "Eligibility requirements", "Can't get time off",  
             "Technical difficulties", "Can't travel to vaccine site",
             "Desired brand not available", "Limited Internet/phone access",
             "Childcare", "Document requirements", "Language barrier"))
 
ggplot(month_barrier_vaccinated_all_long, 
       mapping = aes(x = period_val, y = percentage)) +
      geom_line(aes(color = factor(barriers))) +
      geom_text(data = month_barrier_vaccinated_all_long %>% 
                  group_by(barriers) %>% arrange(desc(period_val)) %>% 
                  slice(1) %>% 
                  filter(percentage >= 35),
            aes(x = period_val - 0.2, y = percentage + 2,
                label = barriers), hjust=0, size=3) +
      geom_vline(xintercept = as.yearmon("202111","%Y%m"), color = "red", linetype = "dashed") +
      geom_text( aes(x = as.yearmon("202111","%Y%m"), y = 0, 
                     label= as.character(as.yearmon("202111","%Y%m"))
                     ), hjust=-0.1, size=3, color = "red") +
      labs(x="month", y="% barriers among all the vaccinated",
           color="Barriers", 
           title="Common barriers to vaccination for the vaccinated",
           subtitle = "From Delphi's surveys, conducted through Facebook",
           caption = "Data from Delphi COVIDcast, delphi.cmu.edu") +
      theme_bw() +
      scale_color_brewer(palette="Paired")
```


Firstly, same as we have noticed in the previous May 2021 table, "No vaccines or
appointments" really stands out from all the other barriers, and the percentage 
is as twice high as the second high barrier "Eligibility requirements". Another 
thing worth noticing is that November 2021 seems to be an important turning 
point for many barriers, which is especially obvious for “No vaccines or 
appointment”. The explanation we have is, on November 19, 2021, [CDC expands eligibility for COVID-19 booster shots to all adults](https://www.cdc.gov/media/releases/2021/s1119-booster-shots.html), so we
were expecting higher demand for vaccination appointments, and thus higher 
barrier percentage.

Since we also ask respondents about their age groups, to have more in-depth idea
on how each age group was affected by different barriers, we decide to divide 
our recipients into 4 age groups: 18-24 years, 25-44 years, 45-64 years and over
65.


```{r}
# specify the factor levels for barriers in the descending order
month_barrier_vaccinated_long$barriers <- factor(
  month_barrier_vaccinated_long$barriers, 
  levels = c("No vaccines or appointments", "Can't get time off",
             "Times didn't work", "Can't travel to vaccine site",
             "Technical difficulties", "Eligibility requirements", 
             "Desired brand not available", "Childcare",
             "Limited Internet/phone access", "Document requirements", 
             "Language barrier"))

ggplot(month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "18-24"),], 
       mapping = aes(x = period_val, y = percentage, color = factor(barriers))) +
      geom_line() +
      geom_text(data = 
          month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "18-24"),] %>% 
                  group_by(barriers) %>% arrange(desc(period_val)) %>% 
                  slice(1) %>% 
                  filter(percentage >= 6),
            aes(x = period_val - 0.16, y = percentage ,
                label = barriers), hjust=0, size=2.2) +
      geom_vline(xintercept = as.yearmon("202111","%Y%m"), color = "red", linetype = "dashed") +
      geom_text( aes(x = as.yearmon("202111","%Y%m"), y = 0, 
                     label= as.character(as.yearmon("202111","%Y%m"))
                     ), hjust=-0.1, size=3, color = "red") +  
      labs(x="month", y="% barriers among all the vaccinated",
           color="Barriers", 
           title="Common barriers to vaccination for the vaccinated aged 18-24",
           subtitle = "From Delphi's surveys, conducted through Facebook",
           caption = "Data from Delphi COVIDcast, delphi.cmu.edu") +
      theme_bw() +
      scale_color_brewer(palette="Paired")
```


For the age group 18-24 years, consistent with the all vaccinated, November 2021
is also a turning point for many barriers, which is obvious for top 3 barriers: 
"no vaccines or appointments", "can’t get time off" and "time didn’t work". 
This is totally understandable, because many in this age group are students or 
hourly workers, and their time is less flexible. To look more closely, both "no 
vaccines or appointments" and "time didn’t work" barriers increase a lot since 
November 2021, thus we suggest add more appointments every time a new 
vaccination is introduced. Since "can’t get time off" basically fluctuates, this
suggests that adding more flexible "vaccination" time-off for hour-based workers
and students should be considered to raise vaccination rate in the long term.


```{r}
# specify the factor levels for barriers in the descending order
month_barrier_vaccinated_long$barriers <- factor(
  month_barrier_vaccinated_long$barriers, 
  levels = c("No vaccines or appointments", "Technical difficulties",
             "Can't travel to vaccine site", "Eligibility requirements",
             "Limited Internet/phone access", "Times didn't work",
             "Desired brand not available", "Can't get time off",
             "Document requirements", "Childcare", "Language barrier"))

ggplot(month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "65plus"),], 
       mapping = aes(x = period_val, y = percentage, color = factor(barriers))) +
      geom_line() +
      geom_text(data = 
          month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "65plus"),] %>% 
                  group_by(barriers) %>% arrange(desc(period_val)) %>% 
                  slice(1) %>% 
                  filter(percentage >= 3),
            aes(x = period_val - 0.1, y = percentage + 0.7 ,
                label = barriers), hjust=0, size=2.2) +
      geom_vline(xintercept = as.yearmon("202111","%Y%m"), color = "red", linetype = "dashed") +
      geom_text( aes(x = as.yearmon("202111","%Y%m"), y = 0, 
                     label= as.character(as.yearmon("202111","%Y%m"))
                     ), hjust=-0.1, size=3, color = "red") +  
      labs(x="month", y="% barriers among all the vaccinated",
           color="Barriers", 
           title="Common barriers to vaccination for the vaccinated aged over 65",
           subtitle = "From Delphi's surveys, conducted through Facebook",
           caption = "Data from Delphi COVIDcast, delphi.cmu.edu") +
      theme_bw() +
      scale_color_brewer(palette="Paired")
```


Also consistent with the all vaccinated, in the age group over 65, November 2021 
is also a turning point for many barriers, indicating the introduce of booster 
also causes effect on the age group over 65. Given the top selected two barriers 
align with what we are expected, this age group will experience trouble making 
appointments online for vaccination. (Another possible explanation for this is 
that the releasing time at the start period of online appointment for vaccination
is usually 12a.m., which is not an ideal time for people who is over 65). Though
"technical difficulties" decreases a lot from May to February, but it still 
remains to be a major problem compared with other barriers. Thus, we suggest 
provide more convenient vaccine appointments/ services for people aged over 65.

As mentioned in the very beginning of the article, cases for the unvaccinated 
can be different from the vaccinated. Since the most barriers of the 
unvaccinated just fluctuate, instead, we will just look at the most recent data 
collected on January 2022 as below:


```{r}
# percentage table for different barriers by vaccination status on May,2021

## sample size without removing "none of above" (barrier 11)
# sample size for the unvaccinated Jan
samplesize_unv1_use <- survey1_full %>% filter(V1==2 & !is.na(V15b)) %>% nrow
samplesize_unv1_1824_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="18-24 years") %>% nrow
samplesize_unv1_2544_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="25-44 years") %>% nrow
samplesize_unv1_4564_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="45-64 years") %>% nrow
samplesize_unv1_65_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="65 years or older") %>% nrow

## sample size removing "none of above" (barrier 11)
# sample size for the unvaccinated Jan
samplesize_unv1 <- jan_v15b_input %>% nrow
samplesize_unv1_1824 <- jan_v15b_input %>% filter(age=="18-24 years") %>% nrow
samplesize_unv1_2544 <- jan_v15b_input %>% filter(age=="25-44 years") %>% nrow
samplesize_unv1_4564 <- jan_v15b_input %>% filter(age=="45-64 years") %>% nrow
samplesize_unv1_65 <- jan_v15b_input %>% filter(age=="65 years or older") %>% nrow

# percentage for barriers who have vaccinated
freq <- table(unlist(jan_v15b_input$barriers))
freq_1824 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="18-24 years")]))
freq_2544 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="25-44 years")]))
freq_4564 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="45-64 years")]))
freq_65 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="65 years or older")]))

prop <- as.data.frame(freq/samplesize_unv1)
prop_1824 <- as.data.frame(freq_1824/samplesize_unv1_1824)
prop_2544 <- as.data.frame(freq_2544/samplesize_unv1_2544)
prop_4564 <- as.data.frame(freq_4564/samplesize_unv1_4564)
prop_65 <- as.data.frame(freq_65/samplesize_unv1_65)

vac_1824 <- as.double(pull(prop_1824, Freq))
vac_2544 <- as.double(pull(prop_2544, Freq))
vac_4564 <- as.double(pull(prop_4564, Freq))
vac_65 <- as.double(pull(prop_65, Freq))

# order the table by frequency of barriers of all unvaccinated in descending order
table_jan <- mutate(prop, vac_1824, vac_2544, vac_4564, vac_65)
table_jan <- table_jan[with(table_jan,order(Freq, decreasing = TRUE)),]


table_jan[,2:6] <- apply(table_jan[,2:6], 2, function(x) percent(x, accuracy=0.01))

colnames(table_jan) <- c("Barriers","All","18-24 years", "25-44 years","45-64 years", "over 65 years")

# transform the barriers abbreviation into long ones
table_jan$Barriers <- lapply(table_jan$Barriers, function(s) unlist(barrier_explain[s]))
```


```{r}
# comparison table on May, 2021 by vaccination status
gt(table_jan[-1,]) %>%
    cols_align(
    align = "left",
    columns = "Barriers"
  ) %>%
    cols_align(
    align = "right",
    columns = c("All","18-24 years", "25-44 years",
                "45-64 years", "over 65 years")
  ) %>%
  tab_header(
    title = md("**Common barriers to vaccination**"),
    subtitle = md("Jan, 2021*")
  ) %>%
    tab_footnote(paste0("Unvaccinated sample size of all, n = ", 
                        prettyNum(samplesize_unv1_use, big.mark = ",")),
                 cells_column_labels("All")) %>%
    tab_footnote(paste0("Unvaccinated sample size of 18-24 years, n = ", 
                        prettyNum(samplesize_unv1_1824_use, big.mark = ",")),
                 cells_column_labels("18-24 years")) %>%
    tab_footnote(paste0("Unvaccinated sample size of 25-44 years, n = ", 
                        prettyNum(samplesize_unv1_2544_use, big.mark = ",")),
                 cells_column_labels("25-44 years")) %>%
    tab_footnote(paste0("Unvaccinated sample size of 45-64 years, n = ", 
                        prettyNum(samplesize_unv1_4564_use, big.mark = ",")),
                 cells_column_labels("45-64 years")) %>%
    tab_footnote(paste0("Unvaccinated sample size of over 65 years, n = ", 
                        prettyNum(samplesize_unv1_65_use, big.mark = ",")),
                 cells_column_labels("over 65 years"))  
```


The picture is consistent across all unvaccinated age groups. Aside from "other"
and "have not tried yet", the top selected barrier for both 18-24 years and 
25-44 years is "can’t get time off", and for both 45-64 years and over 65 years,
it is "can’t travel to vaccination site". In addition to providing more 
available vaccination services for more sites, we also suggest distribute evenly
vaccination supplies of different brands and provide more choices for potential 
vaccine recipients, noting that "desired brands not available" exceed 10% across
all age groups.


## Limitations

As with any scientific study, there are limitations to the results seen above.
Here are a few important things to keep in mind:

* **We're surveying Facebook users.** While we [weight survey
  responses](https://arxiv.org/abs/2009.14675) to ensure their age and gender
  distribution matches the United States population, our respondents do tend to
  be more educated than the national average, which the weights do not correct
  for.
* **The survey is voluntary.** Facebook draws a random sample of active users
  every day and invites them to take the survey via a promotion in their News
  Feed. Many people don't respond to the invitation, and while the
  Facebook-provided survey weights attempt to account for this by using models
  to predict the probability each user will respond, there may still be
  unobserved biases remaining.
* **The sample size gap cannot be ignored.** One result of these biases is that
  our estimate of the percent of Americans who have already received a COVID 
  vaccine is too high for 2021 when compared against official CDC data. Thus, in 
  our May,2021 table, the data should not be very accurate representative of all 
  the American people.
* **Survey responses are simplifications.** Our respondents can select reasons
  for hesitancy from a list, but ticking boxes can't fully represent the
  complexity of their beliefs and the reasons for their lack of trust in COVID
  vaccines. Also, their responses may not always match their behavior: a
  respondent who says they who says they “have not tried yet” get vaccinated may
  decide to get vaccinated when a desired brand dose is offered to them.


## Conclusions

Based on the vaccination barriers of survey responses from hundreds of thousands
of Facebook users, we divide them by two vaccination status and four age groups.
While there is disparity between the vaccinated and the unvaccinated, the most 
common barrier for both is "no vaccines or appointments", and the scenario 
remains the same from May 2021 to February 2022 for all the vaccinated age 
groups. This may suggest we should take extra efforts to ensure access to 
vaccines for potential vaccine recipients. On the top of that, November 2021 
appears to be a turning point, since then, the percentages for many barriers 
have raised. Noting it is when the eligibility for COVID-19 booster shots 
expands to all adults, extra availability for vaccination and service may be 
ensured whenever a new dose is introduced.

To take a closer look, for the vaccinated 18-24 years, "time didn’t work" and 
"can’t get time off" stand high among the barriers, while for the vaccinated 
over 65 years, "technical difficulties" has diminished considerably but remains 
relatively high. There are no magic solutions, but we are suggesting offering 
more "vaccination time-off" for students and time-based workers, and offering 
more convenient vaccine appointments/ services for people aged over 65.

Nonetheless, different from the vaccinated, among all the unvaccinated age 
groups, "desired brands not available" is commonly selected. Thus, we propose 
evenly distribute vaccine supplies between different brands to provide diverse 
options for potential vaccine recipients.

Nonetheless, this demonstrates the value of Delphi's [COVID Symptom
Surveys](https://delphi.cmu.edu/covidcast/surveys/) for studying public
attitudes and behavior during the pandemic. In partnership with Facebook, who
help us distribute the surveys, and the University of Maryland, who run [an
international version](https://covidmap.umd.edu/) in 56 languages worldwide, we
have been able to track the pandemic and its effects on society for over a year
now. And much of our data is easily accessible to researchers, through
aggregates in the [Delphi COVIDcast Epidata
API](https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/fb-survey.html)
and [tables available for
download](https://cmu-delphi.github.io/delphi-epidata/symptom-survey/contingency-tables.html),
and as de-identified individual responses for [researchers who sign Data Use
Agreements](https://cmu-delphi.github.io/delphi-epidata/symptom-survey/data-access.html).
With so many responses daily, we can react quickly and spot local changes. If
you're conducting COVID-related research, we encourage you to explore the data
and see how it can help you.












