---
title: "Common barriers to Covid-19 Vaccination"
author: Naijia Liu
date: 2021-04-23
tags:
  - COVIDcast
  - symptom surveys
  - vaccine
authors:
  - alex
  - robin
  - naijia
heroImage: 
heroImageThumb: 
related:
  - 
summary: |
  Common barriers to vaccination by vaccination status and 
  different age groups.
acknowledgements: |

output:
  html_document:
    code_folding: hide
  blogdown::html_page:
    toc: true
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## This post's filename starts with _ so it is not run by blogdown during the
## post build process, since it depends on private microdata. Instead, we commit
## the HTML.
library(knitr)
library(ComplexUpset)
library(qdapTools)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(gt)
library(xts)
library(directlabels)
library(ggrepel)
library(data.table)
library(readr)
library(scales)
library(anytime)

knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load the fuctions we need
split_options <- function(column) {
  return(strsplit(column, ",", fixed = TRUE))
}

is_selected <- function(vec, selection) {
  selections <- unlist(lapply(
    vec,
    function(resp) {
      if (length(resp) == 0 || all(is.na(resp))) {
        # Qualtrics files code no selection as "" (empty string), which is
        # parsed by `read_csv` as `NA` (missing) by default. Since all our
        # selection items include "None of the above" or similar, treat both no
        # selection ("") or missing (NA) as missing, for generality.
        NA
      } else {
        selection %in% resp
      }
    }))
  return(selections)
}

code_demographics <- function(d) {
  d$race_ethnicity <- case_when(
    d$D6 == 1 ~ "Hispanic",
    d$D7 == 3 ~ "Black",
    d$D7 == 5 ~ "White",
    is.na(d$D7) | is.na(d$D7) ~ NA_character_,
    TRUE ~ "Other"
  )}
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
## data we will be using for the post

# load the data we need for Wave 11 in May
survey5_full <- fread("/Users/bb/Box/CTIS\ data/2021-05-race-ethnicity.csv.gz")


# remove barrier 11 "None of the above"
survey5 <- fread("/Users/bb/Box/CTIS\ data/2021-05-race-ethnicity.csv.gz") %>% 
  filter(is.na(V15a) | V15a != "11",
         is.na(V15b) | V15b != "11")

survey5$barriers_logic <- split_options(coalesce(survey5$V15a, survey5$V15b))


# load the data we need for trend plot
# read in the data
survey_month <- fread("/Users/bb/Box/CTIS\ data/monthly_nation_age.csv.gz")


# load the data we need for Wave 12 in Jan,2022
survey1_full <- fread("/Users/bb/Box/CTIS\ data/2022-01-race-ethnicity.csv.gz")

# new barriers were introduced in Wave 12 in Jan,2022
survey1 <- fread("/Users/bb/Box/CTIS\ data/2022-01-race-ethnicity.csv.gz") %>% 
  filter(is.na(V15c) | V15c != "11",
         is.na(V15b) | V15b != "11")

survey1$barriers_logic <- split_options(coalesce(as.character(survey1$V15c), survey1$V15b))

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Now let's code barriers for May,2021
barriers <- list(
  "1" = "Eligibility requirements",
  "2" = "No vaccines or appointments",
  "3" = "Times didn't work",
  "4" = "Technical difficulties",
  "5" = "Document requirements",
  "6" = "Limited Internet/phone access",
  "7" = "Can't travel to vaccine site",
  "8" = "Language barrier",
  "9" = "Childcare",
  "10" = "Can't get time off",
  "12" = "Desired brand not available",
  "13" = "Have not tried yet"
)

survey5$barriers <- lapply(
  survey5$barriers_logic,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers[s])
    }
  }
)

# extract barrier string from barriers column
survey5$barriers_str <- sapply(
  survey5$barriers,
  function(ss) {
    paste0(ss, collapse = "&")
  }
)

# add column for vaccination status
survey5$vaccinated <- case_when(
  survey5$V1 == 1 ~ "Vaccinated",
  survey5$V1 == 2 ~ "Not vaccinated"
)


# regroup the age into 3 main groups
# "18-34 years", "35-64 years", "65 years or older"
age_logic <- list(
  "1" = "18-24 years",
  "2" = "25-44 years",
  "3" = "25-44 years",
  "4" = "45-64 years",
  "5" = "45-64 years",
  "6" = "65 years or older",
  "7" = "65 years or older")


## extract the subset for May v15a values
# one hot coding for the May v15a values
v15a_input <- mtabulate(survey5$barriers[which(survey5$V1 == 1)])

# add "age" column to the v15a data frame 
v15a_input$age <- lapply(
  as.character(survey5$D2[which(survey5$V1 == 1)]), # for those who are vaccinated
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(age_logic[s])
    }
  }
)

v15a_input$age <- as.character(v15a_input$age)

#remove the "NA" rows for vaccinated
v15a_input <- v15a_input[-which(is.na((survey5$barriers[which(survey5$V1 == 1)]))),]



## extract the subset for May v15b values
# one hot coding for the v15b values
v15b_input <- mtabulate(survey5$barriers[which(survey5$V1 == 2)])

# add "age" column to the v15b data frame 
v15b_input$age <- lapply(
  as.character(survey5$D2[which(survey5$V1 == 2)]),
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(age_logic[s])
    }
  }
)

v15b_input$age <- as.character(v15b_input$age)

#remove the "NA" rows (the highest pike in the plot) for not vaccinated
v15b_input <- v15b_input[-which(is.na((survey5$barriers[which(survey5$V1 == 2)]))),]
```


```{r}
## trend plot for all the vaccinated
# exclude "none" barrier
month_barrier_vaccinated <- data.frame(as.character(survey_month$period_start),
                      survey_month$age,
                      survey_month$val_pct_vaccine_barrier_eligible_has,
                      survey_month$val_pct_vaccine_barrier_no_appointments_has,
                      survey_month$val_pct_vaccine_barrier_appointment_time_has,
                      survey_month$val_pct_vaccine_barrier_technical_difficulties_has,
                      survey_month$val_pct_vaccine_barrier_document_has,
                      survey_month$val_pct_vaccine_barrier_technology_access_has,
                      survey_month$val_pct_vaccine_barrier_travel_has,
                      survey_month$val_pct_vaccine_barrier_language_has,
                      survey_month$val_pct_vaccine_barrier_childcare_has,
                      survey_month$val_pct_vaccine_barrier_time_has,
                      survey_month$val_pct_vaccine_barrier_type_has)

# only take the data from May,2021 to Jan,2021
month_barrier_vaccinated <- month_barrier_vaccinated[21:70,]

colnames(month_barrier_vaccinated) <- c("period_val","age","eligible","no_appointment",
                             "appointment_time","technical_difficultie",
                             "document","technology_access","travel","language",
                             "childcare","time","type")

month_barrier_vaccinated$period_val <- as.Date(month_barrier_vaccinated$period_val,'%Y%m%d')

# aggregate the age groups for all vaccinated
month_barrier_vaccinated_all <- month_barrier_vaccinated %>% group_by(period_val) %>%
  summarise(eligible_all = sum(eligible),
            no_appointment_all = sum(no_appointment),
            appointment_time_all = sum(appointment_time),
            technical_difficultie_all = sum(technical_difficultie),
            document_all = sum(document),
            technology_access_all = sum(technology_access),
            travel_all = sum(travel),
            language_all = sum(language),
            childcare_all = sum(childcare),
            time_all = sum(time),
            type_all = sum(type))

colnames(month_barrier_vaccinated_all) <- colnames(month_barrier_vaccinated)[c(1,3:15)]

# convert the data frame into pivot long table
month_barrier_vaccinated_all_long <- month_barrier_vaccinated_all %>%
  pivot_longer(cols = eligible:type,
               names_to = "barriers",
               values_to = "percentage")

# Now let's code barriers
barriers_name <- list(
  "eligible" = "Eligibility requirements",
  "no_appointment" = "No vaccines or appointments",
  "appointment_time" = "Times didn't work",
  "technical_difficultie" = "Technical difficulties",
  "document" = "Document requirements",
  "technology_access" = "Limited Internet/phone access",
  "travel" = "Can't travel to vaccine site",
  "language" = "Language barrier",
  "childcare" = "Childcare",
  "time" = "Can't get time off",
  "type" = "Desired brand not available"
)

month_barrier_vaccinated_all_long$barriers <- lapply(
  month_barrier_vaccinated_all_long$barriers,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers_name[s])
    }
  }
)

# barriers column is a named list
# convert it into a list
barriers_new <- c()
for (i in 1:length(month_barrier_vaccinated_all_long$barriers)){
  barriers_new <- c(barriers_new,
                    (toString(unname(month_barrier_vaccinated_all_long$barriers)[i])))
}

month_barrier_vaccinated_all_long$barriers <- barriers_new
```


```{r}
## trend plot for the vaccinated by different age groups
# convert the data frame into pivot long table
month_barrier_vaccinated_long <- month_barrier_vaccinated %>%
  pivot_longer(cols = eligible:type,
               names_to = "barriers",
               values_to = "percentage")

month_barrier_vaccinated_long$barriers <- lapply(
  month_barrier_vaccinated_long$barriers,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers_name[s])
    }
  }
)

# barriers column is a named list
# convert it into a list
barriers_new <- c()
for (i in 1:length(month_barrier_vaccinated_long$barriers)){
  barriers_new <- c(barriers_new,
                    (toString(unname(month_barrier_vaccinated_long$barriers)[i])))
}

month_barrier_vaccinated_long$barriers <- barriers_new
```


```{r}
# add "age" column to the Jan,2022 full data frame 
survey1_full$age <- lapply(
  survey1_full$D2,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(age_logic[s])
    }
  }
)

## prepare data for the unvaccinated table for Jan,2022
barriers_jan <- list(
  "1" = "Eligibility requirements",
  "2" = "No vaccines or appointments",
  "3" = "Times didn't work",
  "4" = "Technical difficulties",
  "5" = "Document requirements",
  "6" = "Limited Internet/phone access",
  "7" = "Can't travel to vaccine site",
  "8" = "Language barrier",
  "9" = "Childcare",
  "10" = "Can't get time off",
  "12" = "Desired brand not available",
  "13" = "Have not tried yet",
  "14" = "Available locations did't work",
  "15" = "Other"
)

# data wangling for barriers
survey1$barriers <- lapply(
  survey1$barriers_logic,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(barriers_jan[s])
    }
  }
)

# extract barrier string from barriers column
survey1$barriers_str <- sapply(
  survey1$barriers,
  function(ss) {
    paste0(ss, collapse = "&")
  }
)

# add column for vaccination status
survey1$vaccinated <- case_when(
  survey1$V1 == 1 ~ "Vaccinated",
  survey1$V1 == 2 ~ "Not vaccinated"
)

# add "age" column to the data frame 
survey1$age <- lapply(
  survey1$D2,
  function(s) {
    if (all(is.na(s))) {
      NA
    } else {
      unlist(age_logic[s])
    }
  }
)

# extract the subset for Jan 2022 v15b values
# one hot coding for the Jan 2022 v15b values
jan_v15b_input <- survey1 %>% filter(V1==2 & !is.na(V15b) & V15b!=11) %>%
  select(barriers,age)

```


Since the vaccines first were authorized in December 2020 to healthcare workers, we’ve been trying to vaccinate everyone who is eligible in the US. Early on, vaccines were only available to certain groups; 2 months later, Covid-19 vaccines became generally available across the US, and by May, President Biden announced that they were widely available and everyone should get them.

However, despite all that, many people were still not vaccinated by May, and vaccine uptake has not been as high as initially hoped. And that’s why, through its COVID-19 Trends and Impact Survey, the Delphi Group started to track COVID vaccination barriers to help understand the vaccination barriers by vaccination status and age groups.

Since May 20th, 2021, we have asked all respondents whether they have experienced specific barriers to getting the COVID-19 vaccine based on their vaccination status. We hope, through this article, we can give more insightful ideas and suggestions on why people are unvaccinated and how they can enable helping potential vaccinees get vaccinated.

# Common barriers in May 2021

Let’s start with the common barriers to vaccination on May, 2021 by vaccination status. First, we ask respondents whether they have been vaccinated or not, and according to their vaccination status, they will be given a list of 12 or 13 specific barriers. Respondents can select multiple reasons, so the numbers do not add to 100%. Also, we ask unvaccinated people if they've tried to get vaccinated, and if the answer is no, we don't ask them about barriers. 

This may explain why there are so few unvaccinated people in the table below: many unvaccinated people had not tried to get vaccinated. The huge sample size disparity between the vaccinated and the unvaccinated could possibly affect the accuracy of the measurement of the barriers for all respondents


```{r}
# percentage table for different barriers by vaccination status on May,2021

## sample size without removing "none of above" (barrier 11)
# sample size for the vaccinated May
samplesize_v5 <- survey5_full %>% filter(V1==1 & !is.na(V15a)) %>% nrow
# sample size for the unvaccinated May
samplesize_unv5 <- survey5_full %>% filter(V1==2 & !is.na(V15b)) %>% nrow
# sample size for all May
samplesize_all5 <- samplesize_v5+samplesize_unv5

## sample size removing "none of above" (barrier 11)
## to match with the trend plot
# sample size for the vaccinated May
samplesize_v5_use <- survey5_full %>% filter(V1==1 & !is.na(V15a) & V15a!=11) %>% nrow
# sample size for the unvaccinated May
samplesize_unv5_use <- survey5_full %>% filter(V1==2 & !is.na(V15b) & V15b!=11) %>% nrow
# sample size for all May
samplesize_all5_use <- samplesize_v5_use+samplesize_unv5_use

# percentage for barriers who have vaccinated
freq_vaccinated <- table(unlist(survey5$barriers[which(survey5$V1 == 1)]))
prop_vaccinated <- as.data.frame(freq_vaccinated/samplesize_v5_use)

vac <- as.double(pull(prop_vaccinated, Freq))
vac <- c(vac[1:6],0,vac[7:11]) # no barrier "have not tried yet" for the vaccinated

# percentage for barriers who have not vaccinated
freq_not_vaccinated <- table(unlist(survey5$barriers[which(survey5$V1 == 2)]))
prop_not_vaccinated <- as.data.frame(freq_not_vaccinated/samplesize_unv5_use)

nvac <- pull(prop_not_vaccinated, Freq)

# percentage for barriers for all
freq <- table(unlist(survey5$barriers))
prop <- as.data.frame(freq/(samplesize_v5_use+samplesize_unv5_use))

table <- mutate(prop, vac, nvac)
table <- table[with(table,order(Freq,decreasing = TRUE)),]

# order the table by frequency of barriers of all vaccinated in descending order
table <- table[with(table,order(Freq,decreasing = TRUE)),]
table[,2:4] <- apply(table[,2:4], 2, function(x) percent(x, accuracy=0.01))

colnames(table) <- c("Barriers","All", "Vaccinated","Unvaccinated")

# transform the barriers abbreviation into long ones
new_barriers <- c(
  "There were no vaccines or vaccine appointments available", 
  "I did not meet the eligibility requirements",
  "There were technical difficulties with the website or phone line", 
  "Difficulty traveling to a vaccination site",
  "It was difficult to get time away from work or school", 
  "The available appointment times did not work for me",
  "I could not get the type of vaccine I wanted", 
  "Limited access to internet or phone to schedule an appointment",
  "There is no one to provide childcare while getting the vaccine",
  "I was unable to provide a required document",
  "Information not available in my native language",
  "I have not tried to get the vaccine"
)

table$Barriers <- new_barriers


```


```{r}
# comparison table on May, 2021 by vaccination status
gt(table) %>%
  cols_align(
    align = "left",
    columns = Barriers
  ) %>%
    cols_align(
    align = "right",
    columns = c(All, Vaccinated, Unvaccinated)
  ) %>%
  tab_header(
    title = md("**Common barriers to vaccination**"),
    subtitle = md("*May, 2021*")
  ) %>%
    tab_footnote(paste0("Vaccinated sample size, n = ", 
                        prettyNum(samplesize_v5, big.mark = ",")),
                 cells_column_labels("Vaccinated")) %>%
    tab_footnote(paste0("Unvaccinated sample size, n = ", 
                        prettyNum(samplesize_unv5, big.mark = ",")),
                 cells_column_labels("Unvaccinated")) %>%
    tab_footnote(paste0("All sample size, n = ", 
                        prettyNum(samplesize_all5, big.mark = ",")),
                 cells_column_labels("All"))
```

As you can see, there are similarities and differences between the common barriers to vaccination for the vaccinated and the unvaccinated. “There were no vaccines or vaccine appointments” is the top selected barrier for both. Nevertheless, the other two top reasons commonly selected by the vaccinated are eligibility requirements and technical difficulties, while the unvaccinated most often say the available times didn’t work or they could not get the type of vaccine they wanted, which is unexpected for us. What is noticeable is that “I have not tried yet” barrier also stands high among the unvaccinated. This is what we would expect to happen in normal circumstances: As more people get vaccinated, the people who remain unvaccinated are those who either don’t want to be vaccinated or haven’t yet been able to.

# Changing barriers over time

But these numbers don’t tell the whole story. It’s also very insightful to look at the time trend plot for barriers. The common barriers to vaccination varied throughout the past half year with the announcement for boosters, new policies, the larger population of the vaccinated and so on. In order to have clearer overview on the changes, we plot the trend for the barriers to vaccination for the vaccinated from May, 2021 to February, 2022, and two major discoveries are spotted:

```{r}
# specify the factor levels for barriers in the descending order
month_barrier_vaccinated_all_long$barriers <- factor(
  month_barrier_vaccinated_all_long$barriers, 
  levels = c("No vaccines or appointments", "Times didn't work",
             "Eligibility requirements", "Can't get time off",  
             "Technical difficulties", "Can't travel to vaccine site",
             "Desired brand not available", "Limited Internet/phone access",
             "Childcare", "Document requirements", "Language barrier"))
 
ggplot(month_barrier_vaccinated_all_long, 
       mapping = aes(x = period_val, y = percentage)) +
      geom_line(aes(color = factor(barriers))) +
      scale_x_date(date_breaks = "3 months", date_minor_breaks = "1 month", date_labels = "%b %Y") +  
      geom_text(data = month_barrier_vaccinated_all_long %>% 
                  group_by(barriers) %>% arrange(desc(period_val)) %>% 
                  slice(1) %>% 
                  filter(percentage >= 35),
            aes(x = period_val, y = percentage+2.5,
                label = barriers), hjust= 0.9, size=3) +
      geom_vline(xintercept = as.Date("20211101","%Y%m%d"), color = "red", linetype = "dashed") +
      geom_text( aes(x = as.Date("20211101","%Y%m%d"), y = 0, 
                     label= "Nov 2021"),
                 hjust=-0.1, size=3, color = "red") +
      labs(x="month", y="% barriers among all the vaccinated",
           color="Barriers", 
           title="Common barriers to vaccination for the vaccinated",
           subtitle = "From Delphi's surveys, conducted through Facebook",
           caption = "Data from Delphi COVIDcast, delphi.cmu.edu") +
      theme_bw() +
      scale_color_brewer(palette="Paired")
```


Firstly, same as we have noticed in the previous May 2021 table, "No vaccines or appointments" really stands out from all the other barriers, and the percentage is as twice high as the second high barrier "Eligibility requirements". Another thing worth noticing is that November 2021 seems to be an important turning point for many barriers, which is especially obvious for “No vaccines or 
appointment”. The explanation we have is, on November 19, 2021, [CDC expands eligibility for COVID-19 booster shots to all adults](https://www.cdc.gov/media/releases/2021/s1119-booster-shots.html), so we were expecting higher demand for vaccination appointments, and thus higher barrier percentage.

# Barriers experienced by each age group

Since we also ask respondents about their age groups, to have more in-depth idea on how each age group was affected by different barriers, we decide to divide our recipients into 4 age groups: 18-24 years, 25-44 years, 45-64 years and over 65.


```{r}
# specify the factor levels for barriers in the descending order
month_barrier_vaccinated_long$barriers <- factor(
  month_barrier_vaccinated_long$barriers, 
  levels = c("No vaccines or appointments", "Can't get time off",
             "Times didn't work", "Can't travel to vaccine site",
             "Technical difficulties", "Eligibility requirements", 
             "Desired brand not available", "Childcare",
             "Limited Internet/phone access", "Document requirements", 
             "Language barrier"))

ggplot(month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "18-24"),], 
       mapping = aes(x = period_val, y = percentage, color = factor(barriers))) +
      geom_line() +
      scale_x_date(date_breaks = "3 months", date_minor_breaks = "1 month", date_labels = "%b %Y") +
      geom_text(
        data = month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "18-24"),] %>% 
                  group_by(barriers) %>% arrange(desc(period_val)) %>% 
                  slice(1) %>% 
                  filter(percentage >= 6),
            aes(x = period_val, y = percentage ,
                label = barriers), hjust= 1.1, size=3) +
      geom_vline(xintercept = as.Date("20211101","%Y%m%d"), color = "red", linetype = "dashed") +
      geom_text( aes(x = as.Date("20211101","%Y%m%d"), y = 0, 
                     label= "Nov 2021"), hjust=-0.1, size=3, color = "red") +  
      labs(x="month", y="% barriers among all the vaccinated",
           color="Barriers", 
           title="Common barriers to vaccination for the vaccinated aged 18-24",
           subtitle = "From Delphi's surveys, conducted through Facebook",
           caption = "Data from Delphi COVIDcast, delphi.cmu.edu") +
      theme_bw() +
      scale_color_brewer(palette="Paired")
```


For the age group 18-24, consistent with the all vaccinated, November 2021 is also a turning point for many barriers, which is obvious for top 3 barriers: “no vaccines or appointments”, “can’t get time off” and “time didn’t work”. This is understandable, because many in this age group are students or hourly workers, and their time is less flexible. To look more closely, both “no vaccines or appointments” and “time didn’t work” barriers have increased a lot since November 2021, suggesting it is important to increase vaccination capacity when boosters are introduced. Since 'can't get time off' remains high throughout, this suggests that adding more flexible “vaccination” time-off for hour-based workers and students should be considered to raise vaccination rate in the long term.


```{r}
# specify the factor levels for barriers in the descending order
month_barrier_vaccinated_long$barriers <- factor(
  month_barrier_vaccinated_long$barriers, 
  levels = c("No vaccines or appointments", "Technical difficulties",
             "Can't travel to vaccine site", "Eligibility requirements",
             "Limited Internet/phone access", "Times didn't work",
             "Desired brand not available", "Can't get time off",
             "Document requirements", "Childcare", "Language barrier"))

ggplot(month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "65plus"),], 
       mapping = aes(x = period_val, y = percentage, color = factor(barriers))) +
      geom_line() +
      geom_text(data = 
          month_barrier_vaccinated_long[which(month_barrier_vaccinated_long$age == "65plus"),] %>% 
                  group_by(barriers) %>% arrange(desc(period_val)) %>% 
                  slice(1) %>% 
                  filter(percentage >= 3),
            aes(x = period_val, y = percentage,
                label = barriers), hjust= 1, vjust= -2.2, size= 3) +
      geom_vline(xintercept = as.Date("20211101","%Y%m%d"), color = "red", linetype = "dashed") +
      geom_text( aes(x = as.Date("20211101","%Y%m%d"), y = 0, 
                     label= "Nov 2021"), hjust=-0.1, size=3, color = "red") +  
      labs(x="month", y="% barriers among all the vaccinated",
           color="Barriers", 
           title="Common barriers to vaccination for the vaccinated aged over 65",
           subtitle = "From Delphi's surveys, conducted through Facebook",
           caption = "Data from Delphi COVIDcast, delphi.cmu.edu") +
      theme_bw() +
      scale_color_brewer(palette="Paired")

```

Also consistent with the all vaccinated, in the age group over 65, November 2021 is also a turning point for many barriers, indicating the introduction of boosters also caused changes in the age group over 65. Given the top selected two barriers align with what we expected, this age group will experience trouble making appointments online for vaccination. (Another possible explanation for this is that the releasing time at the start period of online appointment for vaccination is usually 12a.m., which is not an ideal time for people who is over 65). Though “technical difficulties” decreases a lot from May to February, but it still remains to be a major problem compared with other barriers. Thus, we suggest provide more convenient vaccine appointments/ services for people aged over 65.

As mentioned in the very beginning of the article, cases for the unvaccinated can be different from the vaccinated. Since the most barriers of the unvaccinated just fluctuate, instead, look just the most recent data collected on January 2022 as below:


```{r}
## sample size without removing "none of above" (barrier 11)
# sample size for the unvaccinated Jan
samplesize_unv1_use <- survey1_full %>% filter(V1==2 & !is.na(V15b)) %>% nrow
samplesize_unv1_1824_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="18-24 years") %>% nrow
samplesize_unv1_2544_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="25-44 years") %>% nrow
samplesize_unv1_4564_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="45-64 years") %>% nrow
samplesize_unv1_65_use <- survey1_full %>% filter(V1==2 & !is.na(V15b) & age=="65 years or older") %>% nrow

## sample size removing "none of above" (barrier 11)
# sample size for the unvaccinated Jan
samplesize_unv1 <- jan_v15b_input %>% nrow
samplesize_unv1_1824 <- jan_v15b_input %>% filter(age=="18-24 years") %>% nrow
samplesize_unv1_2544 <- jan_v15b_input %>% filter(age=="25-44 years") %>% nrow
samplesize_unv1_4564 <- jan_v15b_input %>% filter(age=="45-64 years") %>% nrow
samplesize_unv1_65 <- jan_v15b_input %>% filter(age=="65 years or older") %>% nrow

# percentage for barriers who have vaccinated
freq <- table(unlist(jan_v15b_input$barriers))
freq_1824 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="18-24 years")]))
freq_2544 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="25-44 years")]))
freq_4564 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="45-64 years")]))
freq_65 <- table(unlist(jan_v15b_input$barriers[which(jan_v15b_input$age=="65 years or older")]))

prop <- as.data.frame(freq/samplesize_unv1)
prop_1824 <- as.data.frame(freq_1824/samplesize_unv1_1824)
prop_2544 <- as.data.frame(freq_2544/samplesize_unv1_2544)
prop_4564 <- as.data.frame(freq_4564/samplesize_unv1_4564)
prop_65 <- as.data.frame(freq_65/samplesize_unv1_65)

vac_1824 <- as.double(pull(prop_1824, Freq))
vac_2544 <- as.double(pull(prop_2544, Freq))
vac_4564 <- as.double(pull(prop_4564, Freq))
vac_65 <- as.double(pull(prop_65, Freq))

# order the table by frequency of barriers of all unvaccinated in descending order
table_jan <- mutate(prop, vac_1824, vac_2544, vac_4564, vac_65)
table_jan <- table_jan[with(table_jan,order(Freq, decreasing = TRUE)),]


table_jan[,2:6] <- apply(table_jan[,2:6], 2, function(x) percent(x, accuracy=0.01))

colnames(table_jan) <- c("Barriers","All","18-24 years", "25-44 years","45-64 years", "over 65 years")


# transform the barriers abbreviation into long ones
new_barriers_jan <- c(
  "Other",
  "There were no vaccines or vaccine appointments available", 
  "The available appointment times did not work for me",
  "It was difficult to get time away from work or school",
  "I have not tried to get the vaccine",
  "Difficulty traveling to a vaccination site",
  "The available appointment locations did not work for me",
  "I could not get the type of vaccine I wanted",
  "I did not meet the eligibility requirements",
  "There were technical difficulties with the website or phone line",
  "Limited access to internet or phone to schedule an appointment",
  "There is no one to provide childcare while getting the vaccine",
  "I was unable to provide a required document",
  "Information not available in my native language"
)

table_jan$Barriers <- new_barriers_jan
```


```{r}
# comparison table on May, 2021 by vaccination status
gt(table_jan) %>%
    cols_align(
    align = "left",
    columns = "Barriers"
  ) %>%
    cols_align(
    align = "right",
    columns = c("All","18-24 years", "25-44 years",
                "45-64 years", "over 65 years")
  ) %>%
  tab_header(
    title = md("**Common barriers to vaccination**"),
    subtitle = md("Jan, 2022*")
  ) %>%
    tab_footnote(paste0("Unvaccinated sample size of all, n = ", 
                        prettyNum(samplesize_unv1_use, big.mark = ",")),
                 cells_column_labels("All")) %>%
    tab_footnote(paste0("Unvaccinated sample size of 18-24 years, n = ", 
                        prettyNum(samplesize_unv1_1824_use, big.mark = ",")),
                 cells_column_labels("18-24 years")) %>%
    tab_footnote(paste0("Unvaccinated sample size of 25-44 years, n = ", 
                        prettyNum(samplesize_unv1_2544_use, big.mark = ",")),
                 cells_column_labels("25-44 years")) %>%
    tab_footnote(paste0("Unvaccinated sample size of 45-64 years, n = ", 
                        prettyNum(samplesize_unv1_4564_use, big.mark = ",")),
                 cells_column_labels("45-64 years")) %>%
    tab_footnote(paste0("Unvaccinated sample size of over 65 years, n = ", 
                        prettyNum(samplesize_unv1_65_use, big.mark = ",")),
                 cells_column_labels("over 65 years"))  
```


The picture is very consistent across all unvaccinated age groups. Aside from “other” and “have not tried yet”, the most common barrier for people aged 18-44 is getting time off of work, while for people aged over 45, it's traveling to the vaccination site. In addition to providing more available vaccination services for more sites, we also suggest distribute evenly vaccination supplies of different brands and provide more choices for potential vaccine recipients, noting that “desired brands not available” exceed 10% across all age groups.


# Limitations

As with any scientific study, there are limitations to the results seen above.
Here are a few important things to keep in mind:

* **We're surveying Facebook users.** While we [weight survey
  responses](https://arxiv.org/abs/2009.14675) to ensure their age and gender
  distribution matches the United States population, our respondents do tend to
  be more educated than the national average, which the weights do not correct
  for.
* **The survey is voluntary.** Facebook draws a random sample of active users
  every day and invites them to take the survey via a promotion in their News
  Feed. Many people don't respond to the invitation, and while the
  Facebook-provided survey weights attempt to account for this by using models
  to predict the probability each user will respond, there may still be
  unobserved biases remaining.
* **The sample size gap cannot be ignored.**The survey overrepresents vaccinated
people. Compared to official CDC data, the survey estimate of the number of 
Americans vaccinated in early 2021 is too high. As a result, the rates of barriers
reported by our respondents may not match the rate experienced by the overall 
American population.
* **Survey responses are simplifications.** Our respondents can select reasons
  for hesitancy from a list, but ticking boxes can't fully represent the
  complexity of their beliefs and the reasons for their lack of trust in COVID
  vaccines. Also, their responses may not always match their behavior: a
  respondent who says they who says they “have not tried yet” get vaccinated may
  decide to get vaccinated when a desired brand dose is offered to them.


# Conclusions

Based on the vaccination barriers of survey responses from hundreds of thousands of Facebook users, we divide them by two vaccination status and four age groups. While there is disparity between the vaccinated and the unvaccinated, the most common barrier for both is no vaccines or appointments available, and the scenario remains the same from May 2021 to February 2022 among all the vaccinated age groups. This may suggest we should take extra efforts to ensure access to vaccines for potential vaccine recipients. On the top of that, November 2021 appears to be a turning point, since then, the percentages for many barriers have raised. Noting it is when the eligibility for COVID-19 booster shots expands to all adults, the demand for vaccination appointments will increase. Extra availability for vaccination and service may need to be be ensured whenever a new dose is introduced.

To take a closer look, for the vaccinated people who are between 18-24 years old, “time didn’t work” and “can’t get time off” stand high among the barriers, while for the vaccinated aged over 65, “technical difficulties” has diminished considerably but remains relatively high. There are no magic solutions, but as [CDC has already suggested](https://www.cdc.gov/vaccines/covid-19/health-departments/generate-vaccinations.html#:~:text=Partner%20with%20local%20community%20groups,or%20request%20at%2Dhome%20vaccinations), we are suggesting offering more “vaccination time-off” for students and time-based workers, and offering more convenient vaccine appointments/ services for priority populations like people aged over 65.

Nonetheless, different from the vaccinated, among all the unvaccinated age groups, the most common barrier is desired brands not available. Thus, we propose evenly distribute vaccine supplies between different brands to provide diverse options for potential vaccine recipients.













